SI

### 2021-02-02 ms/manuscript.Rmd
WIP: add SI figure showing range comparison between G and S
Add light curves to SI

Unlike sporophytes, fern gametophytes are cryptic and cannot be identified to species based on morphology alone. 
Therefore, we used DNA barcoding to identify filmy fern gametophytes to species as described in @Nitta2017.
Briefly, DNA extraction was performed using a modified CTAB protocol [@Doyle1987] or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following @Nitta2017.
For each marker, a reference library was constructed from sporophyte sequences using the 'makeblastDB' command in BLAST [@Altschul1997].
Field-collected gametophytes were identified by a local BLAST query against the reference library.
Gametophytes matching > 99.5 % with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched > 99.5 % with multiple reference sequences were not identified and excluded from further analysis.
For this study, only gametophytes that could be identified as belonging to Hymenophyllaceae were used, and others were excluded from analysis.

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].

@@

Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0.8 s saturating pulse.
This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
500 µmol·m^-2^·s^-1^.
Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
The PPFD at 95 % saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.

@@

In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
All of these have sporophytes that are normally only found in cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.

@@

Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (`r figure("combined")`).
Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
`r mean_par_sporo` µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^.
Similarly, sporophytes attained higher rates of electron transport (ETR~max~; `r mean_etr_sporo` µmol·m^--2^ ·s^--1^) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^).

>>>>

Unlike sporophytes, fern gametophytes are cryptic and cannot be identified to species based on morphology alone. 
Therefore, we used DNA barcoding to identify filmy fern gametophytes to species as described in @Nitta2017.
Briefly, DNA extraction was performed using a modified CTAB protocol [@Doyle1987] or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following @Nitta2017.
For each marker, a reference library was constructed from sporophyte sequences using the \function{makeblastDB} command in BLAST [@Altschul1997].
Field-collected gametophytes were identified by a local BLAST query against the reference library.
Gametophytes matching > 99.5 % with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched > 99.5 % with multiple reference sequences were not identified and excluded from further analysis.
For this study, only gametophytes that could be identified as belonging to Hymenophyllaceae were used, and others were excluded from analysis.

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
Salts used for desiccation and their corresponding mean water potentials and approximate relative humidity are as follows: LiCl (-282 MPa, 18 % RH), Mg(NO~3~)~2~ (-86 MPa, 58 % RH), and NaCl (-38 MPa, 80 % RH).
Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].

@@

Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0.8 s saturating pulse.
This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
500 µmol·m^-2^·s^-1^.
Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is photosynthetic photon flux density (PPFD), *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
The PPFD at 95 % saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.

@@

In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson (`r s_figure("range")`).
All of these have sporophytes that are normally only found in cloud forest (\circa{ }500--1200 m), but gametophytes that are distributed to lower elevations.

@@

Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (\figs `r figure_num("combined")`, `r s_figure_num("light-curves")`).
Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
`r mean_par_sporo` µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^.
Similarly, sporophytes attained higher rates of electron transport (ETR~max~; `r mean_etr_sporo` µmol·m^--2^ ·s^--1^) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^).

### 2021-02-02 ms/si.Rmd
Add light curves to SI

```{r si-setup, include = FALSE, cache = FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
# if running interactively, source _targets.R first
# source("_targets.R")
```

@@

```{r si-load-data, cache = FALSE}
tar_load(filmy_dt_chamber)
tar_load(filmy_dt)
tar_load(filmy_habit)
tar_load(rel_water_indiv)
tar_load(rel_water_species_means)
tar_load(filmy_phy)
tar_load(filmy_habit)
tar_load(range_comparison)
```

@@

@@

}

```

@@

>>>>

```{r si-setup, include = FALSE, cache = FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
# if running interactively, source packages and functions first
# source("R/packages.R")
# source("R/functions.R")
```

@@

```{r si-load-data, cache = FALSE}
tar_load(filmy_dt_chamber)
tar_load(filmy_dt)
tar_load(filmy_habit)
tar_load(rel_water_indiv)
tar_load(rel_water_species_means)
tar_load(filmy_phy)
tar_load(filmy_habit)
tar_load(range_comparison)
tar_load(light_data_all)
tar_load(light_models)
tar_load(filmy_lc_model_params)
tar_load(filmy_lc_model_fitted_data)
```

@@

# Define function to plot facetted light curve
make_facet_lc_plot <- function(light_data, lc_model_fitted_data, lc_model_params) {

  cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
               red = "#D55E00", purple = "#CC79A7")

  ggplot(mapping = aes(x = par)) +
    geom_point(
      data = light_data, 
      aes(y = etr, alpha = outlier)) +
    geom_line(
      data = lc_model_fitted_data, 
      aes(y = etr_fit)) +
    geom_hline(
      data = lc_model_params, 
      aes(yintercept = etr), color = cbPalette[["blue"]]) +
    geom_vline(
      data = lc_model_params, 
      aes(xintercept = par), color = cbPalette[["red"]]) +
    scale_alpha_manual(values = c("TRUE" = 0.2, "FALSE" = 1.0)) +
    labs(
      y = "ETR<sub>max</sub>", 
      x = "PPFD (µmol·m<sup>-2</sup>·s<sup>-1</sup>)", 
      alpha = "Outlier") +
    theme(
      axis.title.x = element_markdown(),
      axis.title.y = element_markdown(),
      strip.text = element_markdown(),
      plot.title = element_markdown(),
      legend.position = "none"
    ) +
    facet_wrap_paginate(
      vars(str_abbrev_sp_pretty(species), light_id), 
      ncol = 4, nrow = 6, page = 1,
      scales = "free", labeller = label_value_2)
}

@@

}
```

@@

```{r sporo-lab-light-curves, fig.height = 9.5, fig.width = 7, fig.cap = ""}
### Make full sets of light curve plots ###
# - sporos in lab
lc_plots_sporo_lab <- make_facet_lc_plot(
  light_data = light_data_all %>% filter(generation == "sporophyte", condition == "lab"),
  lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "sporophyte", condition == "lab"),
  lc_model_params = filmy_lc_model_params %>% filter(generation == "sporophyte", condition == "lab")
)
# - sporos in field
lc_plots_sporo_field <- make_facet_lc_plot(
  light_data = light_data_all %>% filter(generation == "sporophyte", condition == "field"),
  lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "sporophyte", condition == "field"),
  lc_model_params = filmy_lc_model_params %>% filter(generation == "sporophyte", condition == "field")
)
# - gametos
lc_plots_gameto <- make_facet_lc_plot(
  light_data = light_data_all %>% filter(generation == "gametophyte"),
  lc_model_fitted_data = filmy_lc_model_fitted_data %>% filter(generation == "gametophyte"),
  lc_model_params = filmy_lc_model_params %>% filter(generation == "gametophyte")
)

### Get the number of pages needed to print each facetted plot ###
n_sporo_lab_lc_pages <- n_pages(lc_plots_sporo_lab)
n_sporo_field_lc_pages <- n_pages(lc_plots_sporo_field)
n_gameto_lc_pages <- n_pages(lc_plots_gameto)

### Print sporos in the lab ###
# first with title "a"
lc_plots_sporo_lab +
labs(title = "**a**") +
facet_wrap_paginate(
    vars(str_abbrev_sp_pretty(species), light_id),
    ncol = 4, nrow = 6, page = 1,
    scales = "free", labeller = label_value_2)

# print the rest
# with title "a (cont.)"
for(i in 2:n_sporo_lab_lc_pages) {
  p <- lc_plots_sporo_lab +
    labs(title = "**a** (cont.)") +
    facet_wrap_paginate(
      vars(str_abbrev_sp_pretty(species), light_id),
      ncol = 4, nrow = 6, page = i,
      scales = "free", labeller = label_value_2)
  print(p)
}

### Print sporos in the field ###
# first with title "b"
lc_plots_sporo_field +
labs(title = "**b**") +
facet_wrap_paginate(
    vars(str_abbrev_sp_pretty(species), light_id),
    ncol = 4, nrow = 6, page = 1,
    scales = "free", labeller = label_value_2)

# print the rest
# with title "b (cont.)"
for(i in 2:n_sporo_field_lc_pages) {
  p <- lc_plots_sporo_field +
    labs(title = "**b** (cont.)") +
    facet_wrap_paginate(
      vars(str_abbrev_sp_pretty(species), light_id),
      ncol = 4, nrow = 6, page = i,
      scales = "free", labeller = label_value_2)
  print(p)
}

### Print the gametos ###
# first with title "c"
lc_plots_gameto +
labs(title = "**c**") +
facet_wrap_paginate(
    vars(str_abbrev_sp_pretty(species), light_id),
    ncol = 4, nrow = 6, page = 1,
    scales = "free", labeller = label_value_2)

# print the rest
# with title "c (cont.)"
for(i in 2:n_gameto_lc_pages) {
  p <- lc_plots_gameto +
    labs(title = "**c** (cont.)") +
    facet_wrap_paginate(
      vars(str_abbrev_sp_pretty(species), light_id),
      ncol = 4, nrow = 6, page = i,
      scales = "free", labeller = label_value_2)
  print(p)
}
```

**`r s_figure("light-curves")`**. Rapid light response curves of filmy ferns from Moorea, French Polynesia. **a**, sporophytes measured in the lab; **b**, sporophytes measured in the field; **c** gametophytes measured in the field (no gametophytes were measured in the lab). Title of each subplot indicates species followed by light curve ID number assigned to each sample. Light curves fit using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (ETR), *x* is photosynthetic photon flux density (PPFD), *A* is the asymptote of the curve, and *k* is a slope parameter. Outliers at high PPFD not included in the model indicated in light grey. Blue line indicates maximum modeled relative electron transport rate (ETR~max~). Red line indicates PPFD at 95 % saturation of ETR (PPFD~95%~).
